### 2024-02-01 paper.tex
added missing brace
fix typos, one sentence per line

\new{
Overall, VMR quantification yielded results similar to those obtained when quantifying ENCODE regulatory regions, even though the number of detected VMRs (63\,421) is considerably smaller than the number of ENCODE cCREs (339\,815). This demonstrates that our \textit{de novo} VMR detection approach identifies genomic regions in scBS data that are as informative as state-of-the-art epigenomic annotations, as well as its ability to find a parsimonous set of relevant elements. We detected only limited overlap between VMRs and cCREs (Fig.~\ref{fig:score}F), indicating that VMR detection yields information that is complementary to other epigenetic marks.
An added benefit of VMR detection is that this approach is available even in the absence of such annotations, for instance when studying species other than human or mouse.
Furthermore, using VMRs over regulatory regions resulted in decreased RAM requirements as well as a shorter runtime, even when accounting for the additional step of VMR detection.
%Filtering regulatory regions by sequencing coverage reduced resource requirements, but at the cost of a reduced ability to distinguish cell types (Fig.~\ref{fig:resource_usage}).

@@

Here, we have proposed an improved strategy to preprocess single-cell bisulfite sequencing data.
Based on the observation that incomplete read coverage of a genomic interval can lead to inaccurate methylation estimates, we suggest a scoring scheme that is aware of a read's local context rather than just treating all reads in an interval alike. 
\new{In this way our approach offers an alternative to existing tools that impute missing CpG infomation, such as Melissa \citep{kapourani2019melissa}, DeepCpG \citep{angermueller2017deepcpg} or Epiclomal \citep{pe2020epiclomal}. The propper discussion about benefits and drawbacks of data imputation is worth its own paper, but we believe it important to provide the research community with multiple viable options of how to treat their data.}

@@

\new{To furthermore aid interpretation of scBS data, we developed and implemented the first algorithm for detecting dataset-specific DMRs for scBS data.}
FDR estimation via permutation allows us to report statistical significance of each DMR.
\new{In a similar manner as VMRs, pinpointing the regions that differ between groups of cells and hence have a putative regulatory role aids bilogical interpretability.}
\new{For instance,} applying our approach to NSCs and oligodendrocytes demonstrated that the obtained DMRs locate near meaningful loci associated with cell-type specific functions.

We also presented and published an open source software tool, called \texttt{scbs}, that provides an easy-to-use implementation of the described algorithms.
\new{It can start directly from the output of methylation callers such as Bismark \citep{bismark}, Biscuit \citep{biscuit} and methylpy \citep{methylpy} and produce a cell\texttimes region matrix. We suggest iterative PCA as an approach to map the count matrix to a reduced-dimensional space overcoming the abundance of missing values. However, one can also use to the same end established tools based on matrix factorisation such as MOFA+, \citep[][included in our benchmarking]{argelaguet2020mofa} or LIGER \citep{welch2019single}. Once a low dimentional emberdding is obtained, one can switch to well established tools from scRNA-seq analysis, including visualization tools such as t-SNE and UMAP, Leiden/Louvain clustering, pseudotime trajectory analyses, etc.
By offering these functions, \texttt{scbs} bridges a gap in the chain of existing tools that so far hindered practitioners in their data analysis. Our implementation can handle datasets of various sizes up to hundred thousand cells (see Supplementary Figure \ref{fig:umaps}B.}

@@

We note that similar approaches have also been used elsewhere \citep{josse2012}.

\new{Throughout this paper we refer to this approach as iterative PCA.}

>>>>

\new{
Overall, VMR quantification yielded results similar to those obtained when quantifying ENCODE regulatory regions, even though the number of detected VMRs (63\,421) is considerably smaller than the number of ENCODE cCREs (339\,815). This demonstrates that our \textit{de novo} VMR detection approach identifies genomic regions in scBS data that are as informative as state-of-the-art epigenomic annotations, as well as its ability to find a parsimonous set of relevant elements. We detected only limited overlap between VMRs and cCREs (Fig.~\ref{fig:score}F), indicating that VMR detection yields information that is complementary to other epigenetic marks.
An added benefit of VMR detection is that this approach is available even in the absence of such annotations, for instance when studying species other than human or mouse.
Furthermore, using VMRs over regulatory regions resulted in decreased RAM requirements as well as a shorter runtime, even when accounting for the additional step of VMR detection.
%Filtering regulatory regions by sequencing coverage reduced resource requirements, but at the cost of a reduced ability to distinguish cell types (Fig.~\ref{fig:resource_usage}).
}

@@

Here, we have proposed an improved strategy to preprocess single-cell bisulfite sequencing data.
Based on the observation that incomplete read coverage of a genomic interval can lead to inaccurate methylation estimates, we suggest a scoring scheme that is aware of a read's local context rather than just treating all reads in an interval alike. 
\new{In this way our approach offers an alternative to existing tools that impute missing CpG infomation, such as Melissa \citep{kapourani2019melissa}, DeepCpG \citep{angermueller2017deepcpg} or Epiclomal \citep{pe2020epiclomal}.
The proper discussion about benefits and drawbacks of data imputation is worth its own paper, but we believe it important to provide the research community with multiple viable options of how to treat their data.}

@@

\new{To furthermore aid interpretation of scBS data, we developed and implemented the first algorithm for detecting dataset-specific DMRs for scBS data.}
FDR estimation via permutation allows us to report statistical significance of each DMR.
\new{In a similar manner as VMRs, pinpointing the regions that differ between groups of cells and hence have a putative regulatory role aids biological interpretability.}
\new{For instance,} applying our approach to NSCs and oligodendrocytes demonstrated that the obtained DMRs locate near meaningful loci associated with cell-type specific functions.

We also presented and published an open source software tool, called \texttt{scbs}, that provides an easy-to-use implementation of the described algorithms.
\new{It can start directly from the output of methylation callers such as Bismark \citep{bismark}, Biscuit \citep{biscuit} and methylpy \citep{methylpy} and produce a cell\texttimes region matrix.
We suggest iterative PCA as an approach to map the count matrix to a reduced-dimensional space overcoming the abundance of missing values.
However, one can also use to the same end established tools based on matrix factorisation such as MOFA+, \citep[][included in our benchmarking]{argelaguet2020mofa} or LIGER \citep{welch2019single}.
Once a low dimensional embedding is obtained, one can switch to well established tools from scRNA-seq analysis, including visualization tools such as t-SNE and UMAP, Leiden/Louvain clustering, pseudotime trajectory analyses, etc.
By offering these functions, \texttt{scbs} bridges a gap in the chain of existing tools that so far hindered practitioners in their data analysis.
Our implementation can handle datasets of various sizes up to a hundred thousand cells (see Supplementary Figure \ref{fig:umaps}B.}

@@

\new{Throughout this paper we refer to this approach as iterative PCA.}
We note that similar approaches have also been used elsewhere \citep{josse2012}.
